---
title: "SIR Model Example"
# subtitle: "Numerical approximations"
description: "Day 3: Tasks & Activities"
date: 2024-08-07
format: html
toc: true
editor_options: 
  chunk_output_type: console
---

::: {.callout-note}
## **Credit:** 
This example is built on (and closely follows) [this example](http://rstudio-pubs-static.s3.amazonaws.com/6852_c59c5a2e8ea3456abbeb017185de603e.html) presented by Dr. Aidan Findlater
:::

The [**S**usceptible-**I**nfectious-**R**ecovered (SIR) model](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology) is a mathematical model describing the spread of a novel pathogen through a population.

## 1. Load packages
```{r}
#| message: false
#| warning: false
library(tidyverse)
library(deSolve)
```

## 2. Define parameters

```{r}
# Proportion in each compartment: Susceptible 0.999999, Infected 0.000001, Recovered 0 (these are directly from the example linked above) ----
init <- c(S = 1-1e-6, I = 1e-6, R = 0.0)

# beta: infection parameter; gamma: recovery parameter ----
parameters <- c(beta = 2.316, gamma = 0.261)

# time frame ----
times <- seq(0, 100, by = 0.5)
```

## 3. Create the function with the differential equations:

```{r}
# Build the function with all ----
sir <- function(time, init, parameters) {

  with(as.list(c(init, parameters)), {
    dS <- -beta * S * I
    dI <-  beta * S * I - gamma * I
    dR <-  gamma * I
    
    return(list(c(dS, dI, dR)))
    
  })
  
}
```

## 4. Approximate the solution using `deSolve::ode()`:

```{r}
# Solve using `deSolve::ode()` ----
approximation <- ode(y = init, times = times, func = sir, parms = parameters)
```

## 5. Make it something we can plot:
```{r}
# Get output into a data frame ----
approx_df <- as.data.frame(approximation)

# Pivot longer so R will do the work for us ----
approx_long <- approx_df %>% 
  pivot_longer(cols = S:R, names_to = "population", values_to = "proportion")
```

## 6. Plot the output with ggplot:

```{r}
ggplot(data = approx_long, aes(x = time, y = proportion)) +
  geom_line(aes(color = population))
```

